A Study on Common Bug Fix Patterns in Rust

### 2022-11-08 introduction.tex
some changes

In this work, with the help of our tool, Ruxanne, we empirically analyze the most popular Rust projects in GitHub. Our goal is to discover common bug patterns that afflict these projects. Although we are motivated by a desire to develop automated program repair tools, we believe these insights can also be of broader service. Automated program repair~\cite{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit). A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\cite{le2012systematic}. A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\cite{jeffrey2009bugfix}.

>>>>

In this work, with the help of our tool, Ruxanne, we empirically analyze the most popular Rust projects in GitHub. Our goal is to discover common bug patterns that afflict these projects. Although we are motivated by a desire to develop automated program repair tools, we believe these insights can also be of broader service. Automated program repair~\cite{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit). A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\cite{le2012systematic}. A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\cite{jeffrey2009bugfix}. 
Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs. In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust. Ruxanne uses a novel embedding of Rust code into fixed-sized vectors. Through Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories. We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns. We describe each of these patterns and their implications to automated program repair.

### 2022-11-08 methodology.tex
changes to be made
some changes

We decided to implement Ruxanne completely in Python. Our target repositories were the top 18 most-starred open source Rust projects on GitHub as of August 2021. We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:data_modelling} describes how we extracted fixed-size datapoints from the code changes. Section~\ref{sec:mining} describes our mining methodology in depth. Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints. Section~\ref{sec:clustering_data} describes clustering in detail, justifies our choice of DBSCAN for clustering, and explains how we tuned it to improve cluster quality.

@@

The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
\todo[inline]{Worse than that, isn't $d$ unbounded? ``yes it is, but doesn't what we've said above still hold?''}

@@

We propose a novel representation. Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn. However, to simulate order-sensitiveness, we add two more steps. First, we record the number of occurrences of items within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items\todo{is there an example? ``Fig 2?''}. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.

>>>>

We decided to implement Ruxanne completely in Python. Our target repositories were the top 18 most-starred open source Rust projects on GitHub as of August 2021. We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:data_modelling} describes how we extracted fixed-size datapoints from the code changes, and Section~\ref{sec:mining} describes our mining methodology in depth. Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints. Section~\ref{sec:clustering_data} describes clustering in detail, justifies our choice of DBSCAN for clustering, and explains how we tuned it to improve cluster quality.

@@

The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.
\todo[inline]{Worse than that, isn't $d$ unbounded? ``yes it is, but doesn't what we've said above still hold?'' combinatorial explosion}

@@

We propose a novel representation. Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn. However, to simulate order-sensitiveness, we add two more steps. First, we record the number of occurrences of items within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension items\todo{is there an example? ``Fig 2?'' introduction to the next section}. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.
